Optimal Pricing and Informal Sharing: \\ Evidence from Piped Water in Manila

### 2018-10-29 phil_paper/jmp_wjv_v2.tex
some new stuff happening

Despite large investments in piped water throughout the developing world, the share of urban households without piped water has remained stable at 5\% for middle income countries and at 20\% for low income countries for at least the past decade (World Bank [2015]).  Given health, time-savings, and other benefits from piped water, how can water utilities set prices in order to close gaps in access while still covering costs?\footnote{\cite{gamper2010impact,cutler2005role}, and \cite{galiani2005water} find health improvements in both developed and developing settings while \cite{devoto2012happiness} find time savings and other utility benefits.}  A popular approach particularly in developing cities is to subsidize upfront connection fees with high marginal prices to extend access especially for the poor.\footnote{\cite{mcintosh2003asian} and \cite{komives2006distributional} from the World Bank and Asian Development Bank argue for low connection fees.  Using a randomized controlled trial from Morocco, \cite{devoto2012happiness} confirm that modest assistance with connection fees can lead to increases in new water connections.}  In a similar spirit, over 70\% of developing cities increase marginal prices with monthly usage so that large users cross-subsidize access and usage for smaller, poorer users.\footnote{This strategy is also commonly referred to as ``Increasing Block Tariffs'' or  ``Convex Pricing'' (\cite{borenstein2012redistributional}). \cite{hoque2013state} and \cite{boland2000political} provide overviews of the use of increasing marginal prices in the water sector in developing countries.}  Recent empirical papers in South Africa and C{\^o}te d'Ivoire support this intuition, predicting gains in both access to piped water and total welfare from increasing marginal prices (\cite{szabo2015value,diakite2009proposal}).

@@

Let the city of Manila be partitioned into neighborhoods, where each neighborhood is defined by a set of households that are physically close enough to share water connections with each other.  In the context of Manila, these neighborhoods are often characterized by dense clusters of houses or duplexes.  This partition embeds the assumption that neighborhoods are mutually exclusive so that a household located in one neighborhood cannot share water with households in a nearby neighborhood.  Restricting the choice set in this way is necessary for computation but may in some cases overestimate the costs that households face in finding a sharing partner.\footnote{This limitation is discussed in more detail in Section~\ref{section:fixedcostestimation}}

@@

In choosing how much water to use each month, households using piped water face the difficult task of coordinating water consumption among all household members.  For example, consider the head of a household trying to manage water consumption across all of their household members each month.  This task would involve monitoring the meter located outside of their dwelling and coordinating water use among all members.\footnote{Given that many developing countries feature steeply increasing prices with monthly usage, these coordination challenges may magnify any distortions caused by non-linear pricing.}  Consistent with these challenges, mobile survey data indicate that \input{tables/no_info}\unskip\% of surveyed household members do not recall how much they paid for water in the last month while \input{tables/no_use}\unskip\% do not recall the amount that their household consumed in the previous month.

@@

@@

Water sharing relationships impact monthly consumption choices through the budget constraint.\footnote{While the formation of these relationships is discussed in more detail in Section~\ref{section:watersourcechoice}, this section takes these sharing relationships as exogenously given.}  The model assumes that the monthly water bill is split evenly between all sharing households.  Descriptive evidence suggests that challenges observing and coordinating usage within members of a household also exend across households.  Among owners of the water connection responding to the mobile survey, \input{tables/bill_split}\unskip\% report splitting the water bill evenly or alternating payment each month, \input{tables/bill_fixed}\unskip\% charge sharing households a fixed amount each month regardless of their consumption, and \input{tables/bill_usage}\unskip\% charge other users a price according to their usage.\footnote{The remaining households report ``I don't know'' or ``other.''}  Evenly splitting the bill, alternating payment, or charging a fixed amount provide easy ways to extract compensation without requiring specific information on the monthly consumption of each sharing household.  Yet, these methods also introduce free-rider incentives since households do not have to fully pay for their contributions to the total water bill.  Since data limitations prevent identifying the specific type of contract associated with each shared connection, the model assumes that all households evenly split the bill and explores the theoretical and empirical implications of alternate models in Appendix~\ref{appendix:fixedprice}.  The model captures bill splitting by dividing the total monthly payment for a water connection across the $S$ sharing households.

@@

The full budget constraint is a piece-wise expression of equation (\ref{equation:budgetconstraint}) over all $K$ price thresholds.  $B_{C,k}$ represents the total water bill for connection $C$ given the tariff schedule.  This term is calculated by subtracting the current price that the household faces, $p^{k}$, from the previous prices and multiplying by their exposure given by, $(\overline{W}^{j}-\overline{W}^{j-1})$.
%\footnote{In the case where a household uses a connection individually, this expression simplifies by setting the index, $i$, and the total using households, $S$, both equal to one.  In this scenario, the household does not face any hassle costs, $H_i$, and pays the total bill from its corresponding consumption.}

@@

Taking the derivative of expected utility with respect to $w_i$ (holding sharing partners' choices $w_{-i}$ fixed) while letting $\Phi(.)$ equal the standard normal cumulative distribution function produces the following first-order condition:
\begin{align}
\label{equation:firstordercondition}
\begin{split}
 0 =  w_i - \Gamma_i - \alpha_i H_{i} -  \frac{1}{S}  \sum_{j=1}^{K} \alpha_i  p^{j} \Big [ \Phi \Big( \frac{W_C - \overline{W}^{j}}{\sigma_{\epsilon_C}} \Big) - \Phi \Big( \frac{ W_C - \overline{W}^{j-1}}{\sigma_{\epsilon_C}} \Big)   \Big ]
\end{split}
\end{align}
The summation in equation (\ref{equation:firstordercondition}) calculates the expected marginal price faced by the connection, which is scaled by the number of households sharing the connection, $S$.  This expression intuitively captures the extent to which increasing household consumption, $w_i$ --- lifting total consumption, $W_C$ --- exposes the household to higher prices in expectation.  This first-order condition characterizes the best-response functions for each household.
%Higher total consumption, $W_C$, raises the probability that the exogenous shock, $\epsilon_C$, places the households in a higher price range.

To arrive at a tractable expression for water demand, I simplify the best-response function by assuming that households approximate the normal distribution with a simple uniform distribution.  This assumption implies that households give zero likelihood to extremely high or low consumption shocks and even likelihood to moderate consumption shocks.  From an empirical perspective, this function allows for an analytical solution to the demand function while ensuring that the estimation can rely on normally distributed errors to estimate rich preference heterogeneity.

I use the following piece-wise function form to approximate the normal cumulative distribution:

\begin{align*}
%\label{equation:approximation}
\begin{split}
g(z)=& 
\begin{cases}
       0  							  & \text{ if } z    < -\sqrt{2}                       \\
     \frac{1}{2} + \frac{z}{2\sqrt2}  & \text{ if } z \geq -\sqrt{2} \text{ and } z \leq \sqrt{2}  \\
       1 						      & \text{ if } z    >  \sqrt{2}
    \end{cases}
\end{split}
\end{align*}

% This  implies the additional economic assumption that household decision-makers approximate normally distributed shocks with .  
% This approach recovers rich heterogeneity in preferences through a tractable likelihood function.  
% I impose an ex-ante assumption on the size of consumption shocks, $\sigma_{\epsilon_C}$. 
%are partially determined by the variance of the consumption shocks, $\sigma_{\epsilon_C}$.  For example, when $\sigma_{\epsilon_C}$ is very large, demand becomes a function of all prices at every point because the consumption shocks are large enough to produce a positive probability of reaching every price segment regardless of where the households choose to consume.

Appendix~\ref{appendix:linearapprox} graphs this approximation against a standard normal cumulative distribution function and simulates optimal demand under both distributions, finding negligible evidence that this approximation would substantially bias the estimates or affect welfare measures.

After replacing the normal cdf $\Phi(.)$ for $g(.)$, the best-response function in equation (\ref{equation:firstordercondition}) now includes the weighted sum of piece-wise functions, $g(.)$, which may also overlap with each other.  This weighted sum determines the marginal prices that each household must consider in choosing their optimal consumption-level.  The degree of overlap depends on the variance in consumption shocks, $\sigma_{\epsilon_C}$.  For example, when $\sigma_{\epsilon_C}$ is very large, the consumption shocks produce a positive probability reaching any price segment, ensuring full overlap in piece-wise functions.

there is full overlap meaning that households consider all prices at every consumption level.

In the opposite case, having zero consumption shocks produces zero overlap so that households only consider the single price local to their consumption level.

Identifying a particular piece-wise best-response function requires imposing bounds on $\sigma_{\epsilon_C}$ ex-ante, ensuring a certain level of overlap.  I take the approach of 1) assuming bounds on $\sigma_{\epsilon_C}$, 2) constructing the corresponding demand function, and 3) estimating the model to see if the estimated $\sigma_{\epsilon_C}$ lies within the bounds.  Appendix~\ref{appendix:sigmanuassumptions} provides the results from this exercise finding that the results are robust to this ex-ante assumption.

 This section proceeds by describing the demand function that is most consistent with the data.

% determine which prices enter the optimal demand function at each level of demand.

The model also allows for idiosyncratic preference shocks for each household by decomposing the fixed preference of water demand for each household given by $\Gamma_{i}$, into a static, household-specific component, $\gamma_i$, and a normally distributed idiosyncratic component, $\eta_{i}$, which is assumed to follow a multivariate normal distribution among sharing partners:
\begin{align*}
\Gamma_{i} &= \gamma_i + \eta_{i} \\
\eta_{i} &\sim \mathcal{N} (\textbf{0},\Sigma_{\eta}) 
\end{align*}
This component helps to capture unexplained fluctuations in consumption for households over time.  This preference shock is also assumed to be independent of the consumption shocks, $\epsilon_C$.  Aggregating these preferences shocks over all sharing households using connection, $C$, produces a joint shock, $\eta_C$, according to the convolution of these random normal variables with a variance equal to $\sum_{i=1}^{S} \sigma_{\eta}^2 + \sum_{i=1}^{S} \sum_{j>i}^{S} \rho^{\eta}_{i,j} \sigma_{\eta}^2$.

\subsubsection{Characterizing the Optimal Demand Function}

The best-response functions for each household can be solved and summed across all sharing households $S$ to form demand at the level of each connection, $C$.  The optimal demand function takes a piecewise form depending on the proximity of different price segments.\footnote{This approach is only relevant for strictly increasing block tariffs since in this case, the budget constraint is convex.  Tariffs that include both increasing and decreasing block tariffs require comparing indirect utilities across all segments.  I take this calculation into account when allowing for decreasing marginal prices in the counterfactual exercises.}  The aggregate demand function is as follows (where $L,M,H$ index low, medium, and high prices relative to each segment):
\begin{align}\label{equation:generaldemand}
\begin{split}
w^{*}_C &=
\begin{cases}
D^{1}_C   & \text{ if } \hspace{1.7cm}                                               \frac{ \eta_C }{ V_{1,1} }  \leq \overline{\eta}^{1}_{1,L}  \\
D^{1,2}_C   & \text{ if } \hspace{.5cm} \overline{\eta}^{1}_{1,L} <    \frac{ \eta_C }{ V^{1,2} }    \leq \overline{\eta}^{2}_{2,L}  \\
D^{1,2,3}_C   & \text{ if } \hspace{.5cm} \overline{\eta}^{2}_{2,L} <    \frac{ \eta_C }{ V^{1,3} }    \leq \overline{\eta}^{2}_{1,U}  \\
D^{2,3}_C   & \text{ if } \hspace{.5cm} \overline{\eta}^{2}_{1,U} <    \frac{ \eta_C }{ V^{2,3} }    \leq \overline{\eta}^{3}_{2,U}  \\ 
D^{3}_C   & \text{ if } \hspace{.5cm} \overline{\eta}^{3}_{2,U} <    \frac{ \eta_C }{ V^{3,3} }    \leq \overline{\eta}^{3}_{3,L}  \\ 
D^{3,4}_C   & \text{ if } \hspace{.5cm} \overline{\eta}^{3}_{3,L} <    \frac{ \eta_C }{ V^{3,4} }    \leq \overline{\eta}^{4}_{3,U}  \\ 
D^{4}_C   & \text{ if } \hspace{.5cm} \overline{\eta}^{4}_{3,U} <    \frac{ \eta_C }{ V^{4,4} }  
\end{cases}\\
\text{Where: } & \\
D^{L,M,H}_C        		&=  d^{L,M,H}_C + \frac{ \eta_C }{ V_{L,H} }   \\
d^{L,M,H}_C       		&= \frac{ 1}{ V^{L,H} } \Big ( \gamma_C - \alpha_C H_{C} - \frac{\alpha_C}{2 S}  [ p^{H} + p^{L} - \overline{W}^{H} (\frac{p^{H} - p^{M}}{\sqrt{2}{\sigma_{\epsilon_C}}})  - \overline{W}^{M} (\frac{p^{M} - p^{L}}{\sqrt{2}{\sigma_{\epsilon_C}}})  ]  \Big )    \\
V^{L,H} 	     	    &=   1 + \frac{\alpha_C}{2 S} ( \frac{p^{H} - p^{L}}{\sqrt{2}\sigma_{\epsilon_C}})   \\
\overline{\eta}^{j}_{l,L}   &= \overline{W}_l - ( \gamma_C - \alpha_C H_{C} -   \frac{\alpha_C}{S} p^j ) - \sqrt{2} \sigma_{\epsilon_C} \\
\overline{\eta}^{j}_{l,U}   &= \overline{W}_l - ( \gamma_C - \alpha_C H_{C} -   \frac{\alpha_C}{S} p^j ) + \sqrt{2} \sigma_{\epsilon_C} \\
\end{split}
\end{align}

%\begin{align}\label{equation:generaldemand}
%\begin{split}
%w^{*}_C &=
%\begin{cases}
%D^{1}_C   & \text{ if } \hspace{1.7cm}                                               \frac{ \eta_C }{ V_{1,1} }  \leq \overline{\eta}^{1}_{1,L}  \\
%D^{1,2}_C   & \text{ if } \hspace{.5cm} \overline{\eta}^{1}_{1,L} <    \frac{ \eta_C }{ V^{1,2} }    \leq \overline{\eta}^{2}_{2,L}  \\
%D^{1,2,3}_C   & \text{ if } \hspace{.5cm} \overline{\eta}^{2}_{2,L} <    \frac{ \eta_C }{ V^{1,3} }    \leq \overline{\eta}^{2}_{1,U}  \\
%D^{2,3}_C   & \text{ if } \hspace{.5cm} \overline{\eta}^{2}_{1,U} <    \frac{ \eta_C }{ V^{2,3} }    \leq \overline{\eta}^{3}_{2,U}  \\ 
%D^{3}_C   & \text{ if } \hspace{.5cm} \overline{\eta}^{3}_{2,U} <    \frac{ \eta_C }{ V^{3,3} }    \leq \overline{\eta}^{3}_{3,L}  \\ 
%D^{3,4}_C   & \text{ if } \hspace{.5cm} \overline{\eta}^{3}_{3,L} <    \frac{ \eta_C }{ V^{3,4} }    \leq \overline{\eta}^{4}_{3,U}  \\ 
%D^{4}_C   & \text{ if } \hspace{.5cm} \overline{\eta}^{4}_{3,U} <    \frac{ \eta_C }{ V^{4,4} }  
%\end{cases}\\
%\text{Where: } & \\
%D^{L,M,H}_C        		&=  d^{L,M,H}_C + \frac{ \eta_C }{ V_{H,L} }   \\
%d^{L,M,H}_C       		&= \frac{ 1}{ V^{L,H} } \Big ( \gamma_C - \alpha_C H_{C} - \frac{\alpha_C}{2 S}  [ p^{H} (\tfrac{\lambda_C - \overline{W}^{H}}{\lambda_C}) + p^{M} (\tfrac{\overline{W}^{H} - \overline{W}^{H}}{\lambda_C} ) -    p^{L} - \overline{W}^{H} (\frac{p^{H} - p^{M}}{\lambda_C})  - \overline{W}^{M} (\frac{p^{M} - p^{L}}{\lambda_C})  ]  \Big )    \\
%V^{L,H} 	     	    &=   1 + \frac{\alpha_C}{2 S} ( \frac{p^{H} - p^{L}}{\lambda_C})  ,  \lambda_C = \sqrt{2}{\sigma_{\epsilon_C}} \\
%\overline{\eta}^{j}_{l,L}   &= \overline{W}_l - ( \gamma_C - \alpha_C H_{C} -   \frac{\alpha_C}{S} p^j ) - \lambda_C \\
%\overline{\eta}^{j}_{l,U}   &= \overline{W}_l - ( \gamma_C - \alpha_C H_{C} -   \frac{\alpha_C}{S} p^j ) + \lambda_C \\
%\end{split}
%\end{align}

For ease of exposition, I use notation where $D^{l}_C=D^{l,l,l}_C$ and $D^{l,j}_C=D^{l,j,j}_C$.

\node[align=center] at (1.8,1)  { \text{\footnotesize  $D^{1}_C$}};
\node[align=center] at (2.8,1) { \text{\footnotesize  $D^{1,2}_C$}};
\node[align=center] at (3.8,1)  { \text{\footnotesize  $D^{1,2,3}_C $}};
\node[align=center] at (4.8,1) { \text{\footnotesize  $D^{2,3}_C $}};
\node[align=center] at (6.8,1)  { \text{\footnotesize  $D^{3}_C $}};
\node[align=center] at (8.3,1)    { \text{\footnotesize  $D^{3,4}_C $}};
\node[align=center] at (9.6,1)  { \text{\footnotesize  $D^{4}_C $}};

\textbf{ Demand over Many Prices: }  When households choose to consume close to price thresholds points, they become exposed to price segments both above and below the thresholds.  Demand becomes a weighted average of these two price segments.  In this case, demand takes the following form, setting low and medium prices equal to $k$, and the high price equal to $k+1$:
\begin{align*}
D_C &= \frac{ 1}{ V^{k,k+1} } \Big ( \gamma_C - \alpha_C H_{C} - \frac{\alpha_C}{2 S}  [ p^{k} + p^{k+1} - \overline{W}^{k+1} (\frac{p^{k+1} - p^{k}}{\sqrt{2}{\sigma_{\epsilon_C}}})  ]  \Big ) + \frac{ \eta_C }{ V_{k+1,k} } 
\end{align*}
Setting $L=M=k$ simplifies the demand expression so that only the single nearby kink point, $\overline{W}^{k+1}$, is considered.  Under increasing mprices, a decrease in the kink point, $\overline{W}^{k+1}$, lowers demand by increasing the probability that the household faces a higher price.

@@

\sum_{k} f_{\zeta}( Q - D_k ) \Big [ F_{ \Psi | \zeta = Q - D_k } (\overline{\eta}_{T,k}) - F_{ \Psi | \zeta = Q - D_k } (\overline{\eta}_{B,k})  \Big ]

@@

This expression takes the sum of the densities of $\zeta$ for each consumption level weighted by the probability that the preference shock $\Psi$ locates the desired consumption level for that household between the two thresholds, $\overline{\eta}_{T,k}$ and $\overline{\eta}_{B,k}$ over all thresholds.  Given that $\eta_C$ and $\epsilon_C$ are assumed to be independent and normally distributed, the terms in equation (\ref{equation:llcontributionlinear}) can be rewritten using both the convolution and conditional distributions of multivariate normals:

@@

f_{\zeta}( Q - D_k ) &= \frac{1}{\sigma_{\zeta,k}} \phi(\frac{Q - D_k}{\sigma_{\zeta,k}} ) \\

>>>>

Despite large investments in piped water throughout the developing world, the share of urban households without piped water has remained stable at 5\% for middle income countries and at 20\% for low income countries for at least the past decade (World Bank [2015]).  Given health, time-savings, and other benefits from piped water, how can water utilities set prices to ensure access while covering costs?\footnote{\cite{gamper2010impact,cutler2005role}, and \cite{galiani2005water} find health improvements in both developed and developing settings while \cite{devoto2012happiness} find time savings and other utility benefits.}  A popular approach particularly in developing cities is to subsidize upfront connection fees with high marginal prices.\footnote{\cite{mcintosh2003asian} and \cite{komives2006distributional} from the World Bank and Asian Development Bank argue for low connection fees.  Using a randomized controlled trial from Morocco, \cite{devoto2012happiness} confirm that modest assistance with connection fees can lead to increases in new water connections.}  Over 70\% of developing cities also increase marginal prices with monthly usage so that large users cross-subsidize access and usage for smaller, poorer users.\footnote{This strategy is also commonly referred to as ``Increasing Block Tariffs'' or  ``Convex Pricing'' (\cite{borenstein2012redistributional}). \cite{hoque2013state} and \cite{boland2000political} provide overviews of the use of increasing marginal prices in the water sector in developing countries.}  Recent empirical papers in South Africa and C{\^o}te d'Ivoire predict gains in both access and total welfare from increasing marginal prices (\cite{szabo2015value,diakite2009proposal}).

@@

Let the city of Manila be partitioned into neighborhoods, where each neighborhood is defined by a set of households that are physically close enough to share water connections with each other.  In the context of Manila, these neighborhoods are often characterized by dense clusters of houses or duplexes.  This partition embeds the assumption that neighborhoods are mutually exclusive so that a household located in one neighborhood cannot share water with households in a nearby neighborhood.  Restricting the choice set in this way is necessary for computation but may in some cases overestimate the costs that households face in finding a sharing partner.\footnote{This limitation is discussed in more detail in Section~\ref{section:fixedcostestimation}.}

@@

In choosing how much water to consume each month, households using piped water face the difficult task of coordinating water consumption among all household members.  For example, consider the head of a household trying to manage water consumption across all of their household members each month.  This task would involve monitoring the meter located outside of their dwelling and coordinating water use among all members.\footnote{Given that many developing countries feature steeply increasing prices with monthly usage, these coordination challenges may magnify any distortions caused by non-linear pricing.}  Consistent with these challenges, mobile survey data indicate that \input{tables/no_info}\unskip\% of surveyed household members do not recall how much they paid for water in the last month while \input{tables/no_use}\unskip\% do not recall the amount that their household consumed in the previous month.

@@

To account for idiosyncratic preference shocks for water each month, let household's fixed preference, $\Gamma_{i}$, be decomposed into a static, household-specific component, $\gamma_i$, and a normally distributed, idiosyncratic component, $\eta_{i}$, which is assumed to follow a multivariate normal distribution among sharing partners:
\begin{align*}
\Gamma_{i} &= \gamma_i + \eta_{i} \\
\eta_{i} &\sim \mathcal{N} (\textbf{0},\Sigma_{\eta}) 
\end{align*}
This preference shock is also assumed to be independent of the consumption shocks, $\epsilon_C$.  Aggregating these preferences shocks over all sharing households using connection, $C$, produces a joint shock, $\eta_C$, according to the convolution of these random normal variables with a variance equal to $\sum_{i=1}^{S} \sigma_{\eta}^2 + \sum_{i=1}^{S} \sum_{j>i}^{S} \rho^{\eta}_{i,j} \sigma_{\eta}^2$.

@@

Water sharing relationships impact monthly consumption choices through the budget constraint.\footnote{While the formation of these relationships is discussed in more detail in Section~\ref{section:watersourcechoice}, this section takes these sharing relationships as exogenously given.}  The model assumes that the monthly water bill is split evenly between all sharing households.  Descriptive evidence suggests that challenges observing and coordinating usage within members of a household also extend across households.  Among owners of the water connection responding to the mobile survey, \input{tables/bill_split}\unskip\% report splitting the water bill evenly or alternating payment each month, \input{tables/bill_fixed}\unskip\% charge sharing households a fixed amount each month regardless of their consumption, and \input{tables/bill_usage}\unskip\% charge other users a price according to their usage.\footnote{The remaining households report ``I don't know'' or ``other.''}  Evenly splitting the bill, alternating payment, or charging a fixed amount provide easy ways to extract compensation without requiring specific information on the monthly consumption of each sharing household.  Yet, these methods also introduce free-rider incentives since households do not have to fully pay for their contributions to the total water bill.  Since data limitations prevent identifying the specific type of contract associated with each shared connection, the model assumes that all households evenly split the bill and explores the theoretical and empirical implications of alternate models in Appendix~\ref{appendix:fixedprice}.  The model captures bill splitting by dividing the total monthly payment for a water connection across the $S$ sharing households.

@@

The full budget constraint is a piece-wise expansion of equation (\ref{equation:budgetconstraint}) over all $K$ price thresholds.  $B_{C,k}$ represents the total water bill for connection $C$ given the tariff schedule.  This term is calculated by subtracting the current price that the household faces, $p^{k}$, from the previous prices and multiplying by their exposure given by, $(\overline{W}^{j}-\overline{W}^{j-1})$.
%\footnote{In the case where a household uses a connection individually, this expression simplifies by setting the index, $i$, and the total using households, $S$, both equal to one.  In this scenario, the household does not face any hassle costs, $H_i$, and pays the total bill from its corresponding consumption.}

@@

Taking the derivative of expected utility with respect to $w_i$ (holding sharing partners' choices $w_{-i}$ fixed) while letting $\Phi(.)$ equal the standard normal cumulative distribution function produces the following first-order condition:
\begin{align}
\label{equation:firstordercondition}
%\begin{split}
 0 &= - w_i + \Gamma_i - \alpha_i \Big[H_{i} +  \frac{MP_C}{S} \Big]
 \end{align}
\hspace{2cm} Where 
\begin{align}
\label{equation:weightedmarginalprice}
 MP_C &= \sum_{j=1}^{K}  p^{j} \Big [ \Phi \Big( \frac{W_C - \overline{W}^{j}}{\sigma_{\epsilon_C}} \Big) - \Phi \Big( \frac{ W_C - \overline{W}^{j-1}}{\sigma_{\epsilon_C}} \Big)   \Big ]
%\end{split}
\end{align}

The first order condition in equation (\ref{equation:firstordercondition}) reveals that consumers balance their fixed preference for water, $\Gamma_i$, against their price sensitivity, $\alpha_i$, hassle costs, $H_i$, and expected marginal price, $MP_C$ (scaled by the number of sharing households, $S$).

Expected marginal price $MP_C$ in equation (\ref{equation:weightedmarginalprice}) intuitively captures the extent to which increasing household consumption, $w_i$ --- lifting total consumption, $W_C$ --- exposes the household to higher prices in expectation.  This first-order condition characterizes the best-response functions for each household.
%Higher total consumption, $W_C$, raises the probability that the exogenous shock, $\epsilon_C$, places the households in a higher price range.

To arrive at a tractable expression for water demand, I simplify equation (\ref{equation:weightedmarginalprice}) by assuming that households approximate the normal distribution with a simple uniform distribution.  This assumption implies that households give zero likelihood to extremely high or low consumption shocks and even likelihood to moderate consumption shocks.  From an empirical perspective, this function allows for an analytical solution to the demand function while ensuring that the estimation can rely on normally distributed errors to estimate rich preference heterogeneity.

I use the following piece-wise function form to approximate the normal cumulative distribution:
\begin{align*}
%\label{equation:approximation}
\begin{split}
g(z)=& 
\begin{cases}
       0  							  & \text{ if } z    < -\sqrt{2}                       \\
     \frac{1}{2} + \frac{z}{2\sqrt2}  & \text{ if } z \geq -\sqrt{2} \text{ and } z \leq \sqrt{2}  \\
       1 						      & \text{ if } z    >  \sqrt{2}
    \end{cases}
\end{split}
\end{align*}
% This  implies the additional economic assumption that household decision-makers approximate normally distributed shocks with .  
% This approach recovers rich heterogeneity in preferences through a tractable likelihood function.  
% I impose an ex-ante assumption on the size of consumption shocks, $\sigma_{\epsilon_C}$. 
%are partially determined by the variance of the consumption shocks, $\sigma_{\epsilon_C}$.  For example, when $\sigma_{\epsilon_C}$ is very large, demand becomes a function of all prices at every point because the consumption shocks are large enough to produce a positive probability of reaching every price segment regardless of where the households choose to consume.  
Appendix~\ref{appendix:linearapprox} graphs this approximation against a standard normal cumulative distribution function and simulates optimal demand under both distributions, finding negligible evidence that this approximation would substantially bias the estimates or affect welfare measures.

After substituting the normal cumulative distribution function $\Phi(.)$ for $g(.)$ in equation (\ref{equation:weightedmarginalprice}), expected marginal price becomes the sum of potentially overlapping piece-wise functions, $g(.)$.  When households face small consumption shocks, there is no probability that these shocks will overlap and expose the household to multiple price levels.  Therefore, solving the best-response function (summing across all households using connection $C$) yields a simple expression for the deterministic segment of connection-level demand:
\begin{align*}
d^{j}_C &= \gamma_C - \alpha_C [ H_{C} + \frac{p^{j}}{S}]  
\end{align*}
This case produces a linear demand function where $\alpha_C$ can be readily interpreted as the price sensitivity parameter while $\gamma_C$ captures the portion of demand that is invariant to prices.  This demand function captures moral hazard from bill-splitting by dividing the price sensitivity, $\alpha_C$, by the total number of households using the connection, $S$.  Comparing demand among shared connections to individual connections, the extent of over-consumption at an average price of $p$ can be roughly summarized as $\frac{S-1}{S} \alpha_C p $.
%The first condition ensures that the connection's idiosyncratic preference for water, $\eta_C$, falls between the appropriate thresholds, $\overline{\eta}_{L}, \overline{\eta}_{U}$, to be exposed to this particular segment of the price schedule.  The second two conditions ensure that these thresholds do not overlap with each other so that the connection is exposed to a single price for this level of demand.  The thresholds are determined by the kink points, $\overline{W}$, as well as $\sqrt{2} \sigma_{\epsilon_C} $.

%When households face large consumption shocks, these functions overlap across price thresholds, which means that households must consider multiple prices when choosing their consumption.  
This expression for demand holds when total demand (including idiosyncratic preference $\eta_C$) falls sufficiently between two price thresholds: 
\begin{align*}
\overline{W}_{l} + \sqrt{2} \sigma_{\epsilon_C}  \, \leq \, & d^{j}_C + \eta_C \, \leq  \, \overline{W}_{l+1} - \sqrt{2} \sigma_{\epsilon_C} 
\end{align*}
High variances of consumption shocks, $\sigma_{\epsilon_C}$, combined with small distances between kink points create cases where these conditions are not satisfied and households are forced to consider more than one price in choosing optimal consumption.  Allowing for households to be exposed to price thresholds $j$ through $l$, the general solution to the best-response function for the connection $C$ again yields the deterministic segment of connection-level demand as follows:
\begin{align*}
d_{C}^{j,l}       		&= \frac{ 1}{ V^{j,l} } \Big ( \gamma_C - \alpha_C [ H_{C} - \frac{MP^{j,l}}{S} ] \Big )    \\
\text{ Where : }& \\
MP^{j,l} &=\frac{1}{2} \big  [ p^{j} + p^{l} - \sum_{k=l}^{j} \overline{W}^{k} (\frac{p^{k} - p^{k-1}}{\sqrt{2}{\sigma_{\epsilon_C}}}) \big ]  \\
V^{j,l}	     	    &=  1 + \frac{\alpha_C}{2 S} ( \frac{p_j - p_l}{\sqrt{2}\sigma_{\epsilon_C}}) 
\end{align*}
The demand function takes a similar form to the linear case, including the fixed preference while weighting the marginal price by the price sensitivity and degree of sharing.  Unlike in the linear case, connections face an average of all possible marginal prices, $MP^{j,l}$, that can be realized from their consumption shocks.  All terms are also weighted by a variance term, which depends on the gap between the highest and lowest prices $j$ and $l$.  This demand function holds for levels of demand between the following thresholds:
\begin{align*}
B_{j,l} \, & \leq \, D_C^{j,l} \, \leq \, T_{j,l} \\
\text{Where :} & \\
B_{j,l} &= max \{ x \in I_{j,l} \, | \, x \leq D_C^{j,l} \} \,\,\, , \,\,\, T_{j,l} = min  \{ y \in I_{j,l} \, | \, y \geq D_C^{j,l} \} \\
I_{j,l} &= \{ \overline{W}_{l} + \sqrt{2} \sigma_{\epsilon_C}, \overline{W}_{l} - \sqrt{2}\sigma_{\epsilon_C}, \overline{W}_{j} + \sqrt{2} \sigma_{\epsilon_C},  \overline{W}_{j} - \sqrt{2} \sigma_{\epsilon_C}  \} \\
D_C^{j,l} &=  d_C^{j,l} + \frac{ \eta_C }{ V^{j,l} } 
\end{align*}
Bottom $B_{j,l}$ and top $T_{j,l}$ thresholds sandwich connection demand $D_C^{j,l}$ and are drawn from the full set of possible thresholds, $I_{j,l}$, above and below the tariff kink points, $\overline{W}_{l}$ and $\overline{W}_{j}$.  In cases with high variance in consumption shocks, kink points $j$ and $l$ may span many other marginal prices, $k$.  At the same time for small consumption shocks combined with far distances from price thresholds, this expression simplifies to the linear case described above.
%T &= \{ \overline{W}_{0} + \sqrt{2} \sigma_{\epsilon_C}, ..., \overline{W}_{N} + \sqrt{2} \sigma_{\epsilon_C}, \overline{W}_{0} - \sqrt{2} \sigma_{\epsilon_C}, ..., \overline{W}_{N} - \sqrt{2} \sigma_{\epsilon_C}  \} \\
%min \{ \overline{\eta}^{j}_{j,U}, \overline{\eta}^{j}_{l,U} , \overline{\eta}^{j}_{j,L}  , \overline{\eta}^{j}_{l,L} \} \, \leq \, &\frac{ \eta_C }{ V^{j,l} } \, \leq \, max  \{  \overline{\eta}^{j}_{j,U} , \overline{\eta}^{j}_{l,U} , \overline{\eta}^{j}_{j,L}, \overline{\eta}^{j}_{l,L} \} 
%The number of price segments between $j$ and $l$ depends on the extent to which

Constructing a full demand function for a connection requires simply connecting piecewise demand segments to span the full range of consumption choices.  However, since the degree of overlap between price segments depends on the variance in consumption shocks, there can be different optimal demand functions for different levels of consumption variance, $\sigma_{\epsilon_C}$.  To determine a single piecewise demand function, I take the approach of (1) assuming ex-ante bounds on $\sigma_{\epsilon_C}$, (2) constructing the implied optimal demand function given these bounds, and (3) estimating the model to see if the estimated $\sigma_{\epsilon_C}$ rejects these bounds in the data.  Appendix~\ref{appendix:sigmanuassumptions} provides the results of this exercise identifying a range of $\sigma_{\epsilon_C}$ that is consistent with the data.  Appendix~\ref{appendix:optimaldemand} describes the demand function given these bounds along with a visual example demand curve in Figure~\ref{figure:intuitivedemand}.

%Solving for the the best-response function, and summing across all households $C$ yields connection level demand:
%\begin{align*}
%d_{C}^{j,l}       		&= \frac{ 1}{ V^{j,l} } \Big ( \Gamma_C - \alpha_C H_{C} - \frac{\alpha_C}{2S} \big  [ p^{j} + p^%{l} - \sum_{k=l}^{j} \overline{W}^{k} (\frac{p^{k} - p^{k-1}}{\sqrt{2}{\sigma_{\epsilon_C}}}) \big ]  \Big )    \\
%V_{C}^{j,l}	     	    &=   1 + \frac{\alpha_C}{2 S} ( \frac{p_j - p_l}{\sqrt{2}\sigma_{\epsilon_C}}) 
%\end{align*}

@@

\sum_{j} f_{\zeta}( Q - D^{j,l} ) \Big [ F_{ \Psi | \zeta = Q - D^{j,l} } (\overline{\eta}_{T,l}) - F_{ \Psi | \zeta = Q - D^{j,l} } (\overline{\eta}_{B,j})  \Big ]

@@

This expression takes the sum of the densities of $\zeta$ for each consumption level weighted by the probability that the preference shock $\Psi$ locates the desired consumption level for that household between the two thresholds, $\overline{\eta}_{T,l}$ and $\overline{\eta}_{B,j}$ over all thresholds.  Given that $\eta_C$ and $\epsilon_C$ are assumed to be independent and normally distributed, the terms in equation (\ref{equation:llcontributionlinear}) can be rewritten using both the convolution and conditional distributions of multivariate normals:

@@

f_{\zeta}( Q - D^{j,l} ) &= \frac{1}{\sigma_{\zeta}} \phi(\frac{Q - D^{j,l}}{\sigma_{\zeta}} ) \\
